---
title: "PS04"
author: "WRITE YOUR NAME HERE"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

## README

Complete the following questions below and show all work. You must submit a compiled R report with all code and answers in this document. Compile the report to html or PDF and submit on Canvas. html is preferred. Make sure to write your name above where it says “WRITE YOUR NAME HERE”. Include it within the quotation marks.

Before writing answering the following questions, make sure to compile the report to make sure that everything is working properly. As you answer each question, compile after each question to make sure you are not making any errors. If you are experiencing an error, read the error carefully. One tactic is to selectively delete questions/sections until the report compiles. This will help you narrow down where the coding error is occurring.

## Data

The Massachusetts data are district-wide averages for public elementary school districts in 1998. The data analyzed here are the overall total score, which is the sum of the scores on the English, Math, and Science portions of the test. Data on the student-teacher ratio, the percent of students receiving a subsidized lunch and on the percent of students learning english are averages for each elementary school district for the 1997–1998 school year and were obtained from the Massachusetts department of education. Data on average district income are from the 1990 US Census.

## Integrity

If you are suspected of cheating, you will receive a zero—for the assignment and possibly for the course. Cheating includes copying work from your classmates, from the internet, and from previous problem sets. You are encouraged to work with your peers, but everyone must submit their own solutions.

## Questions

### Q01

Load the required packages `tidyverse`, `broom`, and `AER`. Uncomment the code below and compile the document.

```{r}
# Load packages
# library(tidyverse)
# library(broom)

# Load AER package (install if not available)
# if (!require("AER")) install.packages("AER")
```

### Q02

Load data with the following code `data("MASchools")`. Now an object named `MASchools` should be loaded into your R environment. Feel free to rename the object if you wish.

-   Print the first 10 observations in the dataset by using the function `head()`.

```{r}
# data("MASchools")
```

### Q03

Read the documentation of the data by typing `?MASchools`. Describe in words what the following variables mean:

-   `district`
-   `municipality`
-   `stratio`
-   `score4`
-   `exptot`
-   `income`

### Q04

The following functions help users look the data printed on the screen:

-   `str()`
-   `glimpse()`
-   `View()`

Use one or all of these functions to solution the following.

-   What observation level is the data at?
-   What type of data is this (recall data types from the review lecture).

*Aside*: The `skim()` function from the package `skimr` is a good alternative that you may prefer. Though you will need to install and load the package first.

### Q05

Use `dplyr` functions to answer the following questions about the data.

-   How many different school districts/municipalities are there? (_Hint_: Use the `distinct()` function.

```{r}
# Code here
```

-   Which municipality spends the most money on each child?

```{r}
# Code here
```

-   Of the non-missing data, what is the average 4th grade test score?

```{r}
# Code here
```

### Q06

Plot the histogram of test scores for 4th graders. Make it look nice by adding a title and labels to your axis. Maybe try to give it some color. If you would like, install the `ggthemes` package (`install.packages("ggthemes")`) and read the documentation. It has a ton of themes choices that will style your plot.

```{r}
# Code here
```

### Q07

Suppose we are policy makers interested in the relationship between 4th grade test scores and class room size. To save money, we could increase the student to teacher ratio, but it is unclear what the effect on test scores might be. We specify the following model:

$$
\text{Scores}_i = \beta_0 + \beta_1 \text{Class Size}_i + u_i
$$

Before running any regressions,

-   Write down a two tailed hypothesis test for $\beta_1$

$$
H_0: \text{Null Hypothesis Here} \\ 
H_1: \text{Alternative Hypothesis Here}
$$

### Q08

It is always important to plot your data. Make a scatter plot of the variables in this regression. Like before, try to make it look presentable.

```{r}
# Code here
```

### Q09

Run the above regression in R using the `lm` command. Before doing so, read the help file for the `tidy()` and `glance()` functions from the `broom` package.

```{r}
# Code here
```

-   Present the output of the regression as a `tibble` using the `tidy()` function and interpret the estimates.
-   Is the point estimate for $\beta_1$ statistically significant?
-   What is the 95% confidence interval surrounding this estimate?
-   What do we conclude with respect to our earlier hypothesis test?
-   Find the $R^2$ by using the `glance()` function?
-   Does class size explain a lot of the variation in test scores?

### Q10

Describe in math, and in words, what the exogeneity assumption requires. In this context, describe a violation of the exogeneity assumption.

### Q11

What is the formula for omitted variable bias? If we were to include a variable on school funding, what would you expect the sign to be? Are we over or under estimating $\beta_1$? Explain.

### Q12

On your old scatter plot, add another `geom` layer using `geom_smooth`. Read the help file (`?geom_smooth`) to plot the `lm` method, which is the same regression in question Q09. Do you point estimates line up with your fitted line using `geom_smooth`.

### Q13

Suppose we would like to control for school funding using in the following specification:

$$
\text{Scores}_i = \beta_0 + \beta_1 \text{Class Size}_i + \text{Funding}_i + u_i
$$

-   Run this regression, using the variable `expreg` for school funding, and as before, interpret the coefficients and report the $R^2$.
-   Does this regression improve on the $R^2$ found in the SLR above?

### Q14

Based on the change in our point estimate of $\beta_2$ when we add $Funding_i$, what sign is the bias created by omitting this variable? Does this match your expectation from before?

### Q15 

Conduct an F-test to determine if the coefficients on `stratio` and `expreg` are jointly significant at the 1% level. 

$$
H_0: \text{Null Hypothesis Here} \\ 
H_1: \text{Alternative Hypothesis Here}
$$

You must calculate the F-statistic by using the `resid()` function. Recall the formula for the F-statistic is:

$$
F = \frac{(RSS_m - RSS_k) / k}{RSS_k / (n - k - 1)}
$$

where $RSS_m$ is the residual sum of squares from the regression with the added parameters, $RSS_k$ is the residual sum of squares from the original regression, $k$ is the number of added parameters, and $n$ is the number of observations. To determine the critical F-value, use the `qf()` function. 

```{r}
# Calculate the RSS for each model using the residuals

# Calculate the F-statistic with the proper degrees of freedom

# Obtain the critical F-value from the F-distribution
```


